\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-07-11 write_up/Supplementary.tex
Updated text and output latex table of results
Updating supplementary figures
Tidying, discussion updates and additional figure

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../tests/plots/habitat}
	\caption{habitat preference}
	\label{fig:1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/habitat_spwn}
	\caption{spawning habitat preference}
	\label{fig:2}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/Temperature_gradient}
	\caption{Spatiotemporal qemperature gradient}
	\label{fig:3}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/vessel_move_value}
	\caption{movement of a single vessel over a few trips overlaid on the
		value field (i.e. sum of the population abundance x catchability x value}
	\label{fig:13}
\end{figure}

@@

>>>>

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../analysis/habitat}
	\caption{habitat preference}
	\label{fig:1}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../analysis/habitat_spwn}
	\caption{spawning habitat preference}
	\label{fig:2}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/Temperature_gradient}
	\caption{Spatiotemporal temperature gradient}
	\label{fig:3}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{Plots/vessel_move_value}
	\caption{movement of a fleet over a single trip reference overlaid on the
		value field (i.e. sum of the population abundance x catchability x value}
	\label{fig:13}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{./Plots/Matern}
	\caption{Example of different implementation of the Matern correlation
	function on autocorrelation distance}
	\label{fig:16}
\end{figure}

\begin{figure}[!ht]
	\centering
	\includegraphics[width =0.6\linewidth]{./Plots/Closure_fishing_locations_yearly}
	\caption{Closure fishing locations based on annual closures with a
		coarse spatial resolution. Closure location can be seen in red
	in relation to a) before the closure fishing locations, b) after the
	closure fishing locations, c) species 3 habitat distribution}
	\label{fig:17}
\end{figure}

### 2019-07-11 write_up/mixfishsim_ecolmod.tex
Added some text pointers
subheadings
Tidying, discussion updates and additional figure

\usepackage{hhline}
\usepackage{booktabs}
\usepackage[textsize=scriptsize]{todonotes}
%\usepackage[textsize=scriptsize, disable]{todonotes}  %% to turn off notes
\usepackage{eurosym}
\usepackage{longtable}

@@

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time, with varying knowledge of species distributions
and using species non-selective fishing gear. \added[id=PD]{In doing so}
fisheries \deleted[id=PD]{that}catch an assemblage of species
\added[id=PD]{and}\deleted{, known as mixed fisheries} \replaced[id=JJ]{may}{,
	when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} over-quota catch \added[id=JJ]{when managed by
	single species quotas and fishers exhaust one or more quota. This may}
lead to overexploitation of fish populations \added[id=JJ]{\citep{Ulrich2011a,
		Batsleer2015}}.  \added[id=JJ]{Discarding of fish in
	excess of quota limits the ability to maintain fishing mortality within
	sustainable limits \citep{Alverson1994, Crowder1998, Rijnsdorp2007}}
\deleted[id=PD]{; reducing discarding is crucial}and the ability to manage for
the biological and economic sustainability of fisheries\deleted[id=JJ]{ and
	implementation of an ecosystem approach to fisheries}. As such, there
is increasing interest in technical solutions such as gear and spatial closures
as measures to \replaced[id=PD]{reduce unwanted catch}{avoiding
	discard\added[id=JJ]{ing of fish}} \citep{Kennelly2002, Catchpole2008,
	Bellido2011}\textcolor{red}{(ADD COSGROVE REFERENCE HERE)}.  \\

@@

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time\replaced[id=PD]{. Being}{ and
	are} known \added[id=PD]{directly} rather than inferred from sampling
or commercial catch\added[id=PD]{, we can use the population model to validate
	how inference from fisheries-dependent and fisheries independent
	sampling relates to the real population structure in a way we could not
	with real data}. We ii) compare, at different spatial and temporal
aggregations, the simulated population distributions to samples from
fisheries-dependent and fisheries independent catches to test if these are a
true reflection of the relative density of the populations. We then iii)
simulate a fishery closure to protect a species based on different spatial and
temporal data aggregations. We use these evaluations to draw inference on the
utility of commercial data in supporting management decisions.
\todo{\added[id=JJ]{If the paper has two goals this should be clear from the
		start, but may be better over 2 MSs}\added[id=PD]{I would like
		to keep both parts, but have made clearer in how its set out.
		The closure scenarios form validation of the data aggregation,
		rather than effectiveness of the closures themselves - so its a
		continuation of the same question in my eyes}}\deleted[id=PD]{.
	Further, we extend our analysis to a range of spatial and temporal
	scales to assess the impact of these processes on the success of the
	management measure.}\\

@@

\added[id=PD]{The final process results in a population structure and movement
	pattern unique to each species, with population movement occurring on a
	weekly basis. The decision to model population movement on a weekly
	timescale was to reflect that fish tend to aggregate in species
	specific locations and range within a week is fairly limited
	[\textcolor{red}{REF!!}].  Therefore this process approximated the
	demographic shifts in fish populations throughout a year with seasonal
	spawning patterns (e.g.  Figure \ref{S-fig:5}).}

@@

The fleet dynamics can be broadly categorised into three components; fleet
targeting - which determine\replaced[id=PD]{d}{s} the fleet catch efficiency
and preference towards a particular species; trip-level decisions, which
determine\added[id=PD]{d} the initial location to be fished at the beginning of
a trip; and within-trip decisions, determining movement from one fishing spot
to another within a trip. Together, these elements implement an explore-exploit
type strategy for individual vessels to maximise their catch from an unknown
resource distribution (\cite{Bailey2018}). The decision to use an individual
based model for fishing vessels was taken because fishers as a group tend to
show heterogeneity and individual rather than group dynamics
[\textcolor{red}{REF!}]. Therefore in the simulations fleet dynamics are the
productive of individual experiences rather than pre-defined group dynamics.

@@

The following steps are undertaken to determine closures:
\todo{\added[id=JJ]{Procedure unclear. Refer to symbols in methods section or
		switch order starting with description of data type
		etc..}\added[id=PD]{Yes, will redo}}
\begin{enumerate}
	\item Extract data source
	\item Aggregate according to desired spatial and temporal resolution
	\item Interpolate across entire area at desired resolution using simple
		kriging using the \emph{interp} function from the R package
		akima [\textcolor{red}{REF!}].
	\item Close area covering top 5 \% of catch rates 
\end{enumerate}
In total 28 closure scenarios were run which represent combinations of:

@@

In order to answer this question we compare different spatial and temporal
aggregations of the simulated population distributions to:
\begin{enumerate}[label=\alph*)]
	\item \textbf{fisheries-independent data:} the inferred population from
		a fixed-site sampling survey design as commonly used for
		fisheries monitoring purposes;
	\item \textbf{fisheries-dependent data:} the inferred population from
		our fleet model which includes fishery-induced sampling
		dynamics.
\end{enumerate}

Figure \ref{fig:1} shows the aggregated catch composition from each of the data
sources over a ten-year period (to average seasonal patterns) at different
spatial resolutions. The finer spatial grid for the real population (top left)
and commercial data (top middle) show visually similar patterns, though there
are large unsampled areas in the commercial data from a lack of fishing
activity (particularly in the lower left part of the sampling domain). The
survey data at this spatial resolution displays very sparse information about
the spatial distributions of the populations. The slightly aggregated data on a
5 x 5 grid shows similar patterns and, while losing some of the spatial detail,
there remains good consistency between the `real population' and the commercial
data. Survey data starts to pick out some of the similar patterns as the other
data sources, but lacks coverage. The spatial catch information on a 10 x 10
and 20 x 20 grid lose a significant amount of information about the spatial
resolutions for all data sources, and some differences between the survey,
commercial and 'real population' data emerge. \\

Figure \ref{fig:2} shows the consequences of different temporal aggregations of
the data over a ten-year period, with weekly (top), monthly (middle) and yearly
(bottom) catch compositions from across an aggregated 20 x 20 area. As can be
seen by comparison to the 'real population', the monthly aggregation captures
the major patterns seen in the weekly data, albeit missing more subtle
differences. The yearly data results in a constant catch pattern due to the
aggregation process (sometimes known as an aggregation bias). The commercial
data on a weekly basis shows some of the same patterns as the 'real
population', though the first species (in red) is less well represented and
some weeks are missing catches from the area. The monthly data shows some
consistency between the 'real population' and commercial data for species 2 -
4, though species 1 remains under-represented. On an annual basis,
interestingly the commercial data under represents the first species (in red)
while the survey over represents species 1. This is likely due to the biases in
commercial sampling, with the fisheries not targeting the areas where species 1
are present, and the biases in the survey sampling from over representation of
the spatial distribution.

@@

A regression tree (using the R package REEMtree [ref]) highlights that the
factor most contributing to differences in fishing mortality before and after
the closure was the population (72 \% showing that the closures were effective
for population 3), followed by data resolution (21 \%), data type (7 \%) with
the least important factor the timescale ($<$ 1 \%). In general the finer the
spatial resolution of the data used the greater reduction in fishing mortality
for population 3 after the closures (Figure \ref{fig:4}). The notable outliers
are the commercial data at the coarsest spatial resolution (20 x 20) at a
yearly and weekly timescale, where closures were nearly as effective as the
fine-scale resolution. In this case the closures were sufficiently large to
protect the population (and was as effective as when the closure was based on
the "real population"; \textcolor{red}{CHECK THIS, COULD PLOT ACTUAL CLOSURE
	LOCATIONS??}) but this may have consequences in terms of restricting a
much larger area than necessary. \\

@@

Our study evaluates the importance of data scaling and considers potential bias
introduced through data aggregation when using fisheries data to infer
spatio-temporal dynamics of fish populations. Understanding how fishers exploit
multiple heterogeneously distributed fish populations with different catch
limits or conservation status requires detailed understanding of the overlap of
resources; this is difficult to achieve using conventional modelling approaches
due to species targeting in fisheries resulting in preferential sampling
\citep{Martinez-Minaya2018}. Often data are aggregated or extrapolated which
requires assumptions about the spatial and temporal scale of processes. Our
study explores the assumptions behind such aggregation and preferential
sampling to identify potential impacts on management advice. With modern
management approaches increasingly employing more nuanced spatio-temporal
approaches in order to maximise productivity while taking account of both the
biological and human processes operating on different time-frames
\citep{Dunn2016}, understanding assumptions behind the data used -
increasingly a combination of logbook and positional information from vessel
monitoring systems - is vital to ensure measures are effective. \\

We employ a simulation approach to model each of the population and fishery
processes in a hypothetical 'mixed fishery', allowing us to i) evaluate the
consequences of different aggregation assumptions on our understanding of the
spatio-temporal distribution of the underlying fish populations, and ii)
evaluate the effectiveness of a spatial closure given those assumptions. \\

Our approach captures fine scale population and fishery dynamics and their
interaction in a way not usually possible with real data and thus not usually
considered in fisheries simulations. While other simulation frameworks seek to
model individual vessel dynamics based on inferred dynamics from VMS and
logbook records \citep{Bastardie2010}, or as a system to identify measures to
meet particular management goals \citep{Bailey2018}, our framework allows users
to explore the assumptions in modelling observational data and evaluate the
underlying dynamics of such approaches at a fine spatial and temporal scale.
This offers the advantage that larger scale fishery patterns are emergent
properties of the system rather and results can be compared to those obtained
under a statistical modelling framework. \\

COMPARE AGAINST OTHER MODELS: Osmose, Atlantis,... do they use rationale actors
and economic profit making ?

@@

@@

@@

>>>>

\usepackage{hhline}
\usepackage{booktabs}
%\usepackage[textsize=scriptsize]{todonotes}
\usepackage[textsize=scriptsize, disable]{todonotes}  %% to turn off notes
\usepackage{eurosym}
\usepackage{longtable}

@@

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time, with varying knowledge of species distributions
and using species non-selective fishing gear. \added[id=PD]{In doing so}
fisheries \deleted[id=PD]{that}catch an assemblage of species
\added[id=PD]{and}\deleted{, known as mixed fisheries} \replaced[id=JJ]{may}{,
	when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} over-quota catch \added[id=JJ]{when managed by
	single species quotas and fishers exhaust one or more quota. This may}
lead to overexploitation of fish populations \added[id=JJ]{\citep{Ulrich2011a,
		Batsleer2015}}.  \added[id=JJ]{Discarding of fish in
	excess of quota limits the ability to maintain fishing mortality within
	sustainable limits \citep{Alverson1994, Crowder1998, Rijnsdorp2007}}
\deleted[id=PD]{; reducing discarding is crucial}and the ability to manage for
the biological and economic sustainability of fisheries\deleted[id=JJ]{ and
	implementation of an ecosystem approach to fisheries}. As such, there
is increasing interest in technical solutions such as gear and spatial closures
as measures to \replaced[id=PD]{reduce unwanted catch}{avoiding
	discard\added[id=JJ]{ing of fish}} \citep{Kennelly2002, Catchpole2008,
	Bellido2011, Cosgrove2019}.  \\

@@

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time\replaced[id=PD]{. Being}{ and
	are} known \added[id=PD]{directly} rather than inferred from sampling
or commercial catch\added[id=PD]{, we can use the population model to validate
	how inference from fisheries-dependent and fisheries independent
	sampling relates to the real population structure in a way we could not
	with real data}. We ii) compare, at different spatial and temporal
aggregations, the 'real population' distributions to samples from
fisheries-dependent and fisheries independent catches to test if these are a
true reflection of the relative density of the populations. We then iii)
simulate a fishery closure to protect a species based on different spatial and
temporal data aggregations. We use these evaluations to draw inference on the
utility of commercial data in supporting management decisions.
\todo{\added[id=JJ]{If the paper has two goals this should be clear from the
		start, but may be better over 2 MSs}\added[id=PD]{I would like
		to keep both parts, but have made clearer in how its set out.
		The closure scenarios form validation of the data aggregation,
		rather than effectiveness of the closures themselves - so its a
		continuation of the same question in my eyes}}\deleted[id=PD]{.
	Further, we extend our analysis to a range of spatial and temporal
	scales to assess the impact of these processes on the success of the
	management measure.}\\

@@

\added[id=PD]{The final process results in a population structure and movement
	pattern unique to each species, with population movement occurring on a
	weekly basis. The decision to model population movement on a weekly
	timescale was to reflect that fish tend to aggregate in species
	specific locations which have been observed to last around one to two
	weeks \citep{Poos2007}.  Therefore this process approximated the
	demographic shifts in fish populations throughout a year with seasonal
	spawning patterns (e.g.  Figure \ref{S-fig:5}).}

@@

The fleet dynamics can be broadly categorised into three components; fleet
targeting - which determine\replaced[id=PD]{d}{s} the fleet catch efficiency
and preference towards a particular species; trip-level decisions, which
determine\added[id=PD]{d} the initial location to be fished at the beginning of
a trip; and within-trip decisions, determining movement from one fishing spot
to another within a trip. Together, these elements implement an explore-exploit
type strategy for individual vessels to maximise their catch from an unknown
resource distribution \cite{Bailey2018}. The decision to use an individual
based model for fishing vessels was taken because fishers are heterogeneous in
their location choice behaviour due to different objectives, risk preference
and targeting preference \citep{VanPutten2012a}. Therefore in the simulations
fleet dynamics are the productive of individual experiences rather than
pre-defined group dynamics.

@@

The following steps are undertaken to determine closures:
\todo{\added[id=JJ]{Procedure unclear. Refer to symbols in methods section or
		switch order starting with description of data type
		etc..}\added[id=PD]{Yes, will redo}}
\begin{enumerate}
	\item Extract data source
	\item Aggregate according to desired spatial and temporal resolution
	\item Interpolate across entire area at desired resolution using simple
		kriging using the \emph{interp} function from the R package
		akima \citep{Akima2006}.
	\item Close area covering top 5 \% of catch rates 
\end{enumerate}
In total 28 closure scenarios were run which represent combinations of:

@@

In order to answer this question we compare different spatial and temporal
aggregations of the 'real population' distributions to:
\begin{enumerate}[label=\alph*)]
	\item \textbf{fisheries-independent data:} the inferred population from
		a fixed-site sampling survey design as commonly used for
		fisheries monitoring purposes;
	\item \textbf{fisheries-dependent data:} the inferred population from
		our fleet model which includes fishery-induced sampling
		dynamics.
\end{enumerate}

Figure \ref{fig:1} shows the aggregated catch composition from each of the data
sources over a ten-year period (to average seasonal patterns) at different
spatial resolutions. The finer spatial grid for the real population (top left)
and commercial data (top middle) show visually similar patterns, though there
are large unsampled areas in the commercial data from a lack of fishing
activity (particularly in the lower left part of the sampling domain). The
survey data at this spatial resolution displays very sparse information about
the spatial distributions of the populations. The slightly aggregated data on a
5 x 5 grid shows similar patterns and, while losing some of the spatial detail,
there remains good consistency between the `real population' and the commercial
data. Survey data starts to pick out some of the similar patterns as the other
data sources, but lacks spatiotemporal coverage. The spatial catch information
on a 10 x 10 and 20 x 20 grid lose a significant amount of information about
the spatial resolutions for all data sources, and some differences between the
survey, commercial and 'real population' data emerge. \\

Figure \ref{fig:2} shows the consequences of different temporal aggregations of
the data over a ten-year period, with weekly (top), monthly (middle) and yearly
(bottom) catch compositions from across an aggregated 20 x 20 area. By
comparison to the 'real population', the monthly aggregation captures the major
patterns seen in the weekly data, albeit missing more subtle differences. The
yearly data results in a constant catch pattern due to the aggregation process
(sometimes known as an aggregation bias). The commercial data on a weekly basis
shows some of the same patterns as the 'real population', though the first
species (in red) is less well represented and some weeks are missing catches
from the area. The monthly data shows some consistency between the 'real
population' and commercial data for species 2 - 4, though species 1 remains
under-represented. On an annual basis, interestingly the commercial data under
represents the first species (in red) while the survey over represents species
1. This is likely due to the biases in
commercial sampling, with the fisheries not targeting the areas where species 1
are present, and the biases in the survey sampling from over representation of
the spatial distribution.

@@

A regression tree (using the R package REEMtree \citep{Sela2012}) highlights
that the factor most contributing to differences in fishing mortality before
and after the closure was the population (72 \% showing that the closures were
effective for population 3), followed by data resolution (21 \%), data type (7
\%) with the least important factor the timescale ($<$ 1 \%). In general the
finer the spatial resolution of the data used the greater reduction in fishing
mortality for population 3 after the closures (Figure \ref{fig:4}). The notable
outliers are the commercial data at the coarsest spatial resolution (20 x 20)
at a yearly and weekly timescale, where closures were nearly as effective as
the fine-scale resolution. In this case the closures were sufficiently large to
protect a core are of the habitat for the population, but this was achieved in
a fairly crude manner by closing a large area - including area where the
species was not found (Figure \ref{S-fig:17}) which may have consequences in
terms of restricting the fishery in a much larger area than necessary. \\

@@

%% General intro

Our study evaluates the importance of data scaling and considers potential bias
introduced through data aggregation when using fisheries data to infer
spatiotemporal dynamics of fish populations. Understanding how fishers exploit
multiple heterogeneously distributed fish populations with different catch
limits or conservation status requires detailed understanding of the overlap of
resources; this is difficult to achieve using conventional modelling approaches
due to species targeting in fisheries resulting in preferential sampling
\citep{Martinez-Minaya2018}. Often data are aggregated or extrapolated which
requires assumptions about the spatial and temporal scale of processes. Our
study explores the assumptions behind such aggregation and preferential
sampling to identify potential impacts on management advice. With modern
management approaches increasingly employing more nuanced spatiotemporal
approaches in order to maximise productivity while taking account of both the
biological and human processes operating on different time-frames
\citep{Dunn2016}, understanding assumptions behind the data used - increasingly
a combination of logbook and positional information from vessel monitoring
systems - is vital to ensure measures are effective. \\

We employ a simulation approach to model each of the population and fishery
dynamics in a hypothetical 'mixed fishery', allowing us to i) evaluate the
consequences of different aggregation assumptions on our understanding of the
spatiotemporal distribution of the underlying fish populations, and ii)
evaluate the effectiveness of a spatial closure given those assumptions. \\

%% USP
Our approach is unique in that it captures fine scale population and fishery
dynamics and their interaction in a way not usually possible with real data and
thus not usually considered in fisheries simulations. While other simulation
frameworks seek to model individual vessel dynamics based on inferred dynamics
from VMS and logbook records \citep{Bastardie2010}, or as a system to identify
measures to meet particular management goals \citep{Bailey2018}, our framework
allows users to explore the assumptions in modelling observational data and
evaluate the underlying dynamics of such approaches at a fine spatial and
temporal scale.  This offers the advantage that larger scale fishery patterns
are emergent properties of the system rather and results can be compared to
those obtained under a statistical modelling framework. \\

%% Comparison to other models
Typically, simulation models which treat fish as individuals are focussed on
exploring the inter- and intra- specific interactions among fish populations
(e.g. OSMOSE \cite{Shin2004}) in order to understand how they vary over space
and time. Our focus was on understanding how fleets which exploit multiple fish
populations realise catch distributions which may differ from the underlying
populations. As such, we favoured a minimum realistic model of the fish
populations after \cite{Plaganyi2014}, while incorporating fishing dynamics
which take account of different drivers in a mechanistic way, similarly to
other approaches \citep{Fulton2011} but at an individual vessel rather than
fleet level.

%% Why ours is important

@@

%% key findings in context

@@

\subsection{Future application for framework}

@@

\subsection{Model assumptions and caveats}
